Supplementary Information: Vulnerability of tree growth

### 2025-03-25 doc/manuscript_hkk_drought_sensitivity.Rmd
removed conflict

<!--PARAGRAPH ON IMPORTANCE OF DROUGHT, HOW OUR UNDERSTANDING REMAINS LIMITED, PARTICULARLY FOR SEASONAL TROPICAL FORESTS-->

Increasing drought risk is threatening demographic processes in global forests and consequently their consistent contributions to the land carbon sink [IPCC].  
- Every year, woody growth of trees in tropical forests removes roughly `r round (150*1/6,0)` Gt CO~2~ from the atmosphere [@beer_terrestrial_2010;@anderson-teixeira_carbon_2021;@anderson-teixeira_carbon_2016], contributing to a net biomass increase of ~7 Gt CO~2-eq~ [@harris_global_2021] that offsets ~`r round(7/36.3*100,-1)`% of anthropogenic fossil fuel emissions [@friedlingstein_global_2022].<!--copied from our grant proposal - something like this is more specific and perhaps more compelling-->
- Moreover, woody growth has one of the longest residence times in terrestrial carbon pool, contributing to long-lasting impacts on carbon removal from the atmosphere [@yang_variations_2021; @muller-landau_patterns_2021].  
- However, tree growth is vulnerable to drought, the frequency and intensity of which are increasing across the much of tropics [@tao_increasing_2022]. <!--and there is some evidence that stand-level vulnerability is increasing [@luo_tropical_2022; @tao_increasing_2022].-->
- (We don't know very much about woody growth responses, as few tropical species form rings and long-term annual records are rare [@zuidema_tropical_2022].)
- Drought sensitivity is often understood as a species trait, but it can vary substantially among drought events [] and among individuals based on their characteristics [@bennett_larger_2015; @mcgregor_tree_2021], compounding uncertainties in predicting forest responses.  
- With changing climate patterns across the global tropics, emergence of novel climatic regimes like "very extreme droughts" in recent decades [@dahinden_future_2017; @luo_tropical_2022], there is an urgent need to understand the drivers of this variation, to be able to predict and respond to forest drought responses.

>>>>

<!--PARAGRAPH ON IMPORTANCE OF DROUGHT, HOW OUR UNDERSTANDING REMAINS LIMITED, PARTICULARLY FOR SEASONAL TROPICAL FORESTS-->
Increasing drought risk is threatening demographic processes in global forests and consequently their consistent contributions to the land carbon sink [IPCC].  
- Every year, woody growth of trees in tropical forests removes roughly `r round (150*1/6,0)` Gt CO~2~ from the atmosphere [@beer_terrestrial_2010;@anderson-teixeira_carbon_2021;@anderson-teixeira_carbon_2016], contributing to a net biomass increase of ~7 Gt CO~2-eq~ [@harris_global_2021] that offsets ~`r round(7/36.3*100,-1)`% of anthropogenic fossil fuel emissions [@friedlingstein_global_2022].<!--copied from our grant proposal - something like this is more specific and perhaps more compelling-->
- Moreover, woody growth has one of the longest residence times in terrestrial carbon pool, contributing to long-lasting impacts on carbon removal from the atmosphere [@yang_variations_2021; @muller-landau_patterns_2021].  
- However, tree growth is vulnerable to drought, the frequency and intensity of which are increasing across the much of tropics [@tao_increasing_2022]. <!--and there is some evidence that stand-level vulnerability is increasing [@luo_tropical_2022; @tao_increasing_2022].-->
- (We don't know very much about woody growth responses, as few tropical species form rings and long-term annual records are rare [@zuidema_tropical_2022].)
- Drought sensitivity is often understood as a species trait, but it can vary substantially among drought events [] and among individuals based on their characteristics [@bennett_larger_2015; @mcgregor_tree_2021], compounding uncertainties in predicting forest responses.  
- With changing climate patterns across the global tropics, emergence of novel climatic regimes like "very extreme droughts" in recent decades [@dahinden_future_2017; @luo_tropical_2022], there is an urgent need to understand the drivers of this variation, to be able to predict and respond to forest drought responses.
